---
title: "Implementation of ridreg regression Class `ridreg` "
author: "Saman and Rabnawaz"
header-includes:
   - \usepackage{statPack}
package: statPack
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Using the caret package and your ridgereg() function to create a predictive model for the BostonHousing data found in the mlbench package.

## The Document should include the following:
- Divide the BostonHousing data (or your own API data) into a test and training dataset using the
caret package.
- Fit a linear regression model and a linear regression model with forward selection of covariates on the training dataset. Information on linear regression models in the caret package can be found here http://topepo.github.io/caret/Linear Regression.html.
- Evaluate the performance of this model on the training dataset.
- Fit a ridge regression model using your ridgereg() function to the training dataset for different
values of. How to include custom models in caret is described here
http://topepo.github.io/caret/custom models.html.
- Find the best hyperparameter value for using 10-fold cross-validation on the training set. More
information how to use the caret package for training can be found here
https://cran.r-project.org/web/packages/caret/vignettes/caret.pdf and here
http://topepo.github.io/caret/training.html.
- Evaluate the performance of all three models on the test dataset and write some concluding com-
ments.

### 1. Divide the BostonHousing data
```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(caret)
library(mlbench)
library(statPack)

data(BostonHousing)

```


### Creating Test and Training Data set
```{r}
data("BostonHousing") #load a data
boston_data <- BostonHousing #set a data to variable
indexes = createDataPartition(boston_data$medv, p = .75, list = FALSE, times = 1)
training<- boston_data[indexes,] #assigninng 75% data to test
testing<- boston_data[-indexes,]  #assigning remaining 25% data to training set

```

The data has now been divided into a training and a test data set.

## Linear regression and model evaluation

### lm method

A linear regression model on the training function can be fitted with the `train()` function from the caret package.

### 2. Fit Linear Regression Model and Linear Regression Model with Forward selection

```{r}
set.seed(-312312L)
ridgereg_fit <- train(rm ~ . , data = training, method = "lm")
print(ridgereg_fit)

ridgereg_forward_fit <- train(rm ~ ., data = training, method = "leapForward")
print(ridgereg_forward_fit)
```
### 3. performance of this model on the training dataset.
The lm is better value f RMSE and MAE than leap forward so `Lm` is better



The first model with the 'lm' method has a better RMSE and MAE value which indicates a better performance with the first model.

### 4. Creating Custom Model for Ridge Regression
```{r}
ridge <- list(type="Regression", 
              library="statPack",
              loop=NULL,
              prob=NULL)
ridge$parameters <- data.frame(parameter="lambda",
                               class="numeric",
                               label="lambda")
ridge$grid <- function(y,x, len=NULL, search="grid"){
  data.frame(lambda=c(0.1,0.5,1,2))
}
ridge$fit <- function (x, y, wts, param, lev, last, classProbs, ...) {
  dat <- if (is.data.frame(x)) 
    x
  else as.data.frame(x)
  dat$.outcome <- y
  out <- ridgereg$new(.outcome ~ ., data = dat, lambda=param$lambda, ...)
  out
}
ridge$predict <- function (modelFit, newdata, submodels = NULL) {
  if (!is.data.frame(newdata)) 
    newdata <- as.data.frame(newdata)
  newdata[,apply(newdata, MARGIN=2, sd)!=0] <- scale(newdata[,apply(newdata, MARGIN=2, sd)!=0])
  modelFit$predict(newdata)
}
```

```{r}
#result will store in train function
result <- train( medv ~ ., data=training, method=ridge)
```


### 5. Appication of 10-fold cross validation
```{r}
fitControl <- control <- trainControl(method = "repeatedcv",
                        number=10,
                        repeats = 10)

 result <- train(crim ~ ., data = training,method = ridge,preProc = c("scale","center"),
       tuneLength = 10,trControl = fitControl)
````


### 6.Evaluation of Models:
Based on the RMSE values of each model, it is estimated that `linear model` is better than `ridereg and leap forward regressions`.

## `Repo link`

> "Here you can find a private repo link which will be public soon"
([Rcourse-Lab7](https://github.com/rjkhan/RCourse-lab7))

#### query
`rabnsh696@student.liu.se` or `samza595@student.liu.se` 